---
title: "EMBL-EBI SARS-CoV-2 data-ingress"
output: 
    flexdashboard::flex_dashboard:
        theme: flatly
        social: menu
        orientation: rows
        vertical_layout: fill
        storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(curl)
library(tidyverse)
library(maps)
library(viridis)
library(ggrepel)
library(ggstance)
#library(ggalt)
library(highcharter)
library(lubridate)
library('rjson')
library(silgelib)
theme_set(theme_plex())

options(warn=-1)

query<-  "https://www.ebi.ac.uk/ena/portal/api/search?result=read_experiment&query=tax_tree(2697049)%20AND%20experiment_accession%3D%22ERX*%22&fields=center_name,country,scientific_name,instrument_platform,library_layout,instrument_model,collection_date,first_public,sex,host&format=tsv"


curl_download(query, 'metadata.tsv')
metadata <- readr::read_delim(file='metadata.tsv',delim = "\t")


# NB: For new center_name to country mapping
# follow the instruction in the comment section of the code below.
metadata_clean <- metadata %>% transmute(
  origin = case_when(
    str_detect(center_name, fixed('aalborg',ignore_case = TRUE)) ~ 'Denmark',
    str_detect(center_name, fixed('deepseq',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('fisabio',ignore_case = TRUE)) ~ 'Spain',
    str_detect(center_name, fixed('DR. RICARDO JORGE',ignore_case = TRUE)) ~ 'Portugal',
    str_detect(center_name, fixed('Colindale',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Cardiff',ignore_case = TRUE)) ~ 'Wales',
    str_detect(center_name, fixed('Quadram Institute Bioscience',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('University College London',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('University of Cambridge',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('University of Edinburgh',ignore_case = TRUE)) ~ 'Scotland',
    str_detect(center_name, fixed('University of Glasgow',ignore_case = TRUE)) ~ 'Scotland',
    str_detect(center_name, fixed('Liverpool Clinical Laboratories',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('University of Sheffield',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Belfast Health',ignore_case = TRUE)) ~ 'Northern Ireland',
    str_detect(center_name, fixed('University of Birmingham',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('University of Exeter',ignore_case = TRUE)) ~ 'England',
    # FOR NEW CENTER_NAME to COUNTRY MAPPING, un-comment  and fill in the following template
    #str_detect(center_name, fixed('CENTER_NAME',ignore_case = TRUE)) ~ 'COUNTRY'),
    TRUE ~ ifelse (!is.na(country), country, center_name)),
  country, experiment_accession, sample_accession, center_name, scientific_name,
  instrument_platform, library_layout, instrument_model, collection_date, first_public, sex, host
) %>%
  mutate_if(is.character, factor)


# few european countries
eu <- c(
  "Portugal", "Spain", "France", "Switzerland", "Germany",
  "Austria", "Belgium", "UK", "Netherlands",
  "Denmark", "Norway","Finland","Sweden","Israel","Ireland","Malta","Poland", "Italy",
  "Croatia", "Slovenia", "Hungary", "Slovakia",
  "Czech republic"
)
# Retrieve the map data
eu.map <- map_data("world", region = eu)
world.map <- map_data("world")

# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
eu.lab <- eu.map %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))


uk_states <- eu.map%>% filter(str_detect('UK', region)) %>%
  group_by(subregion) %>%
  summarize(long=mean(long), lat=mean(lat)) %>%
  rename(region=subregion)
uk_states <- bind_rows(uk_states ,tibble(data.frame(region='England', lat=52, long=0.12)))

eu_countries <- eu.map %>% filter(!str_detect('UK', region)) %>%
  group_by(region) %>%
  summarize(long=mean(long), lat=mean(lat))

geo_loc <- bind_rows(uk_states, eu_countries)

meta_df <- metadata_clean %>% count(origin, sort=TRUE) %>% rename(region=origin)

hits <- inner_join(geo_loc, meta_df)

world.center  <- world.map %>%  group_by(region) %>%
        summarize(long=mean(long), lat=mean(lat))
world.center <- bind_rows(world.center, uk_states) 
whits <- inner_join(world.center, meta_df)
```


# Maps {data-icon="fa-map-marker"} 

Row
-----------------------------------------------------------------------

### Total Experiments {.value-box}

```{r}
exp_count <- metadata_clean %>% select(experiment_accession) %>% distinct %>% count() %>% as.numeric()
valueBox(exp_count, icon="ion-ios-flask-outline")
```

### Submission centers {.value-box}

```{r}
center_count <- metadata_clean %>% select(center_name) %>% distinct %>% count() %>% as.numeric()
valueBox(center_count, icon="ion-ios-people-outline")
```


### Instrument models 

```{r}
models_count <- metadata_clean %>% select(instrument_model) %>% distinct ()%>% count() %>% as.numeric()
valueBox(models_count, icon="ion-paintbucket")
```



### Instrument Platforms 

```{r}
platform_count <- metadata_clean %>% select(instrument_platform) %>% distinct ()%>% count() %>% as.numeric()
valueBox(platform_count, icon="ion-paintbucket")
```


## Row  {.tabset .tabset-fade}

### Europe


```{r, fig.height=7, fig.width=9}
mybreak=seq(0,nrow(metadata_clean),2000)
mylabel=paste(gsub("000","",mybreak),"K", sep="")
ggplot() + 
  geom_map(data=eu.map, map=eu.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  geom_text(aes(x=long, y=lat, label = region), data = geo_loc,  size = 2, hjust = 0.5, family="Palatino") +
  geom_label_repel(data=hits, aes(x=long, y=lat ,label=n), size=2, alpha=.9, box.padding = 0.5,colour='slateblue', show.legend=TRUE)+
  stat_summary_hex(
    data=hits,
    aes(long, lat, z=as.integer(n)),
    fun ="mean",
    alpha = 0.7, 
    bins=80
  ) +
  #scale_fill_gradient(high="cyan3",labels=scales::comma) +
  scale_fill_gradient(high="cyan3", breaks=mybreak, labels=mylabel) +
  theme_void(base_family = "IBMPlexSans") + 
  labs(x=NULL, y=NULL, fill=NULL) + 
  theme(plot.title= element_text(hjust=.5), 
        legend.position = "bottom")

```


### World

```{r}
source('fileexport.R')
data(worldgeojson, package = "highcharter")
subset_data <-  metadata_clean %>% mutate(origin=as.character(origin), country=as.character(country)) %>%
    mutate(country = ifelse(origin=="Scotland", "United Kingdom", country)) %>% 
    mutate (country = ifelse(origin=="England", "United Kingdom", country))%>%
    mutate (country = ifelse(origin=="Wales", "United Kingdom", country))%>%
    mutate(country = ifelse(origin =="Spain", origin, country)) %>%
    mutate(country=as.factor(country)) %>%
    count(country) %>% filter(!is.na(country))

hc <- highchart() %>%
  hc_add_series_map(
    worldgeojson , subset_data , value = "n", joinBy = c('name','country'),
    name = "CovidSubmission"
    )  %>% 
  hc_colorAxis(stops = color_stops()) %>%
      hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
hc
```


### Europe2


```{r}
source('fileexport.R')
json_file='https://code.highcharts.com/mapdata/custom/european-union.geo.json'
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

subset_data <-  metadata_clean %>% mutate(origin=as.character(origin), country=as.character(country)) %>%
    mutate(country = ifelse(origin=="Scotland", "United Kingdom", country)) %>% 
    mutate (country = ifelse(origin=="England", "United Kingdom", country))%>%
    mutate (country = ifelse(origin=="Wales", "United Kingdom", country))%>%
    mutate(country = ifelse(origin =="Spain", origin, country)) %>%
    mutate(country=as.factor(country)) %>%
    count(country) %>% filter(!is.na(country))

hc <- highchart() %>%
  hc_add_series_map(
    json_data , subset_data , value = "n", joinBy = c('name','country'),
    name = "CovidSubmission"
    )  %>% 
  hc_colorAxis(stops = color_stops()) %>%
      hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
hc
```



### World2

```{r, fig.height=7, fig.width=9}

ggplot() + 
  geom_map(data=world.map, map=world.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  geom_label_repel(data=whits, aes(x=long, y=lat ,label=n), alpha=.9, box.padding = 0.5, colour='slateblue', show.legend=TRUE)+
  geom_point(data=whits, aes(x=long, y=lat ,size=n), size=2, alpha=.6, colour='red')+
  scale_fill_gradient(high="cyan3", labels= scales::comma) +
  theme_void(base_family = "IBMPlexSans") + 
  labs(x=NULL, y=NULL, fill="Number of\n samples")  + 
  theme(legend.position = "none")


```



# Center Names {data-icon="ion-ios-people-outline"} 

## Row {.tabset .tabset-fade}

### Submission Centers

```{r, fig.height=7, fig.width=9}
source('fileexport.R')

metadata_clean %>% count(center_name, origin) %>% arrange(desc(n)) %>% mutate(center_name =fct_reorder(center_name, n)) %>%
  hchart("bar", hcaes(x=center_name , y=n, group=origin)) %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
```


# Timeline {data-icon="fa-calendar"} 

## Row {.tabset .tabset-fade}

### Overall

```{r, fig.height=7, fig.width=9}
source('fileexport.R')

metadata_clean %>% count(first_public, origin) %>% hchart("column", hcaes(x=first_public , y=n, group=origin)) %>%
    hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
```

### Month

```{r, fig.height=7, fig.width=9}
metadata_clean %>% 
    mutate(monthly=floor_date(first_public, unit="month")) %>%
    count(monthly, origin) %>% 
    hchart("column", hcaes(x= monthly, y=n, group=origin))%>%
    hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))

```


### Week

```{r, fig.height=7, fig.width=9}

metadata_clean %>% 
    mutate(weekly=week(first_public)) %>%
    count(weekly, origin) %>% 
    hchart("column", hcaes(x= weekly, y=n, group=origin))%>%
    hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))


```

Platforms {data-icon="ion-erlenmeyer-flask"}
=====================================     

## Row {.tabset .tabset-fade}

### Instrument model 

```{r}

metadata_clean %>% count(instrument_model, instrument_platform) %>% mutate(instrument_model = fct_reorder(instrument_model, n)) %>% ggplot(aes(instrument_model, n, fill=instrument_platform)) + geom_col() + coord_flip() 
```

### Instrument platforms 

```{r}

metadata_clean %>% count(instrument_model, instrument_platform, library_layout) %>% mutate(instrument_model = fct_reorder(instrument_model, n)) %>% ggplot(aes(instrument_model, n, fill=instrument_platform)) + geom_col( show.legend = FALSE) + coord_flip() + facet_grid(library_layout~instrument_platform, scales="free_x")
```



Datasets {data-icon="ion-ios-cloud-download-outline"}
=====================================     

## DT::datatable {.tabset .tabset-fade}

### Basic

```{r}
library(DT)
datatable(metadata_clean)
```          

### Advanced

```{r}

datatable(metadata_clean, filter = 'top',  
          options = list(   searchHighlight = TRUE, pageLength = 15 ))

```


### Exporting

```{r}
library(DT)

metadata_clean %>% datatable( extensions = 'Buttons',  
                              options = list(  dom = 'Bfrtip',
                                               buttons = list( list(
                                               extend = 'collection',
                                        buttons = list(list(extend='csv',filename='embl-ebi-covid-submission.csv'),                                                                 list(extend='excel', filename = 'embl-ebi-covid-submission.xlsx')),text="Download"
                                            ))))

```

