---
title: "EMBL-EBI SARS-CoV-2 data-ingress"
output: 
    flexdashboard::flex_dashboard:
        theme: flatly
        social: menu
        orientation: rows
        vertical_layout: fill
        storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(curl)
library(tidyverse)
library(maps)
library(viridis)
library(ggrepel)
library(ggstance)
#library(ggalt)
library(highcharter)
library(lubridate)
library('rjson')
library(silgelib)
theme_set(theme_plex())

options(warn=-1)

# Query ENA API
experiment<-  "https://www.ebi.ac.uk/ena/portal/api/search?result=read_experiment&query=tax_tree(2697049)&fields=center_name,country,scientific_name,instrument_platform,library_layout,instrument_model,collection_date,first_public,sex,host&format=tsv"

sequence_release_full <- "https://www.ebi.ac.uk/ena/portal/api/search?result=sequence_release&query=tax_tree(2697049)&fields=accession,study_accession,sample_accession,tax_id,scientific_name,base_count,bio_material,cell_line,cell_type,collected_by,collection_date,country,cultivar,culture_collection,dataclass,description,dev_stage,ecotype,environmental_sample,first_public,germline,host,identified_by,isolate,isolation_source,keywords,lab_host,last_updated,location,mating_type,mol_type,organelle,serotype,serovar,sex,submitted_sex,specimen_voucher,strain,sub_species,sub_strain,tax_division,tissue_lib,tissue_type,topology,variety,altitude,haplotype,plasmid,sequence_md5&format=tsv"


sequence_update_full <- "https://www.ebi.ac.uk/ena/portal/api/search?result=sequence_update&query=tax_tree(2697049)&fields=accession,study_accession,sample_accession,tax_id,scientific_name,base_count,bio_material,cell_line,cell_type,collected_by,collection_date,country,cultivar,culture_collection,dataclass,description,dev_stage,ecotype,environmental_sample,first_public,germline,host,identified_by,isolate,isolation_source,keywords,lab_host,last_updated,location,mating_type,mol_type,organelle,serotype,serovar,sex,submitted_sex,specimen_voucher,strain,sub_species,sub_strain,tax_division,tissue_lib,tissue_type,topology,variety,altitude,haplotype,plasmid,sequence_md5&format=tsv"

sequence_release <- "https://www.ebi.ac.uk/ena/portal/api/search?result=sequence_release&query=tax_tree(2697049)&fields=accession,study_accession,sample_accession,tax_id,scientific_name,collected_by,collection_date,country,first_public,last_updated,location&format=tsv"


sequence_update <- "https://www.ebi.ac.uk/ena/portal/api/search?result=sequence_update&query=tax_tree(2697049)&fields=accession,study_accession,sample_accession,tax_id,scientific_name,collected_by,collection_date,country,first_public,last_updated,location&format=tsv"


# Download metadata
curl_download(experiment, 'experiment.tsv')
curl_download(sequence_release, 'seq_release.tsv')
curl_download(sequence_update, 'seq_update.tsv')
# Read in the metadata
experiment <- readr::read_delim(file='experiment.tsv',delim = "\t")
seqrelease <- readr::read_delim(file='seq_release.tsv',delim = "\t")
sequpdate <- readr::read_delim(file='seq_update.tsv',delim = "\t")

seqrelease <- seqrelease %>% mutate(country=gsub(":.*","", country)) %>%
  select(accession, study_accession, sample_accession,country, first_public, last_updated ) %>% 
  mutate(INSDC = case_when(
    str_detect(study_accession,fixed('PRJEB', ignore_case = TRUE)) ~ 'EBI', 
    str_detect(study_accession, fixed('PRJNA',ignore_case = TRUE))~'NCBI', 
    str_detect(study_accession, fixed('PRJDA', ignore_case=TRUE))~'DDBJ')) %>%
  mutate(INSDC = case_when(
    str_detect(sample_accession,fixed('SAMEA', ignore_case = TRUE)) ~ 'EBI', 
    str_detect(sample_accession, fixed('SAMN',ignore_case = TRUE))~'NCBI', 
    str_detect(sample_accession, fixed('SAMD', ignore_case=TRUE))~'DDBJ',
    TRUE ~ INSDC )) %>%
   mutate(
    country = case_when(
    str_detect(country, fixed('USA',ignore_case = TRUE)) ~ 'United States',
    TRUE ~ country
  )) 


sequpdate <- sequpdate %>% mutate(country=gsub(":.*","", country)) %>%
  select(accession, study_accession, sample_accession,country, first_public, last_updated ) %>% 
  mutate(INSDC = case_when(
    str_detect(study_accession,fixed('PRJEB', ignore_case = TRUE)) ~ 'EBI', 
    str_detect(study_accession, fixed('PRJNA',ignore_case = TRUE))~'NCBI', 
    str_detect(study_accession, fixed('PRJDA', ignore_case=TRUE))~'DDBJ')) %>%
  mutate(INSDC = case_when(
    str_detect(sample_accession,fixed('SAMEA', ignore_case = TRUE)) ~ 'EBI', 
    str_detect(sample_accession, fixed('SAMN',ignore_case = TRUE))~'NCBI', 
    str_detect(sample_accession, fixed('SAMD', ignore_case=TRUE))~'DDBJ',
    TRUE ~ INSDC )) %>%
   mutate(
    country = case_when(
    str_detect(country, fixed('USA',ignore_case = TRUE)) ~ 'United States',
    TRUE ~ country
  )) 


# NB: For new center_name to country mapping
# follow the instruction in the comment section of the code below.
experiment_clean <- experiment %>% mutate(country=gsub(":.*","", country)) %>% transmute(
  origin = case_when(
    str_detect(center_name, fixed('aalborg',ignore_case = TRUE)) ~ 'Denmark',
    str_detect(center_name, fixed('deepseq',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('fisabio',ignore_case = TRUE)) ~ 'Spain',
    str_detect(center_name, fixed('DR. RICARDO JORGE',ignore_case = TRUE)) ~ 'Portugal',
    str_detect(center_name, fixed('Colindale',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Wales',ignore_case = TRUE)) ~ 'Wales',
    str_detect(center_name, fixed('Quadram Institute Bioscience',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('London',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Cambridge',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Edinburgh',ignore_case = TRUE)) ~ 'Scotland',
    str_detect(center_name, fixed('Glasgow',ignore_case = TRUE)) ~ 'Scotland',
    str_detect(center_name, fixed('Liverpool',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Sheffield',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Belfast Health',ignore_case = TRUE)) ~ 'Northern Ireland',
    str_detect(center_name, fixed('Birmingham',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Exeter',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Portsmouth',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('EMBL-EBI',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Northumbria',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('Cumbria',ignore_case = TRUE)) ~ 'England',
    str_detect(center_name, fixed('SC',ignore_case = FALSE)) ~ 'England',
    center_name %in% c('SUB7478134','SUB7604810','SUB7121476','SUB7373126')  ~ 'United States',
    # FOR NEW CENTER_NAME to COUNTRY MAPPING, un-comment  and fill in the following template
    #str_detect(center_name, fixed('CENTER_NAME',ignore_case = TRUE)) ~ 'COUNTRY'),
    TRUE ~ ifelse (!is.na(country), country, center_name)),
  country, experiment_accession, sample_accession, center_name, scientific_name,
  instrument_platform, library_layout, instrument_model, collection_date, first_public, sex, host
) %>%
  mutate(insdc= case_when(
    str_detect(experiment_accession, fixed('ERX', ignore_case = TRUE)) ~ "EBI",
    str_detect(experiment_accession, fixed('SRX', ignore_case = TRUE)) ~ "NCBI",   
    str_detect(experiment_accession, fixed('DRX', ignore_case = TRUE)) ~ "DDBJ",
    TRUE ~ 'Others'
  )) %>%
  mutate(
    country = case_when(
    str_detect(country, fixed('USA',ignore_case = TRUE)) ~ 'United States',
    TRUE ~ country
  ),
  origin =  case_when(
    str_detect(origin, fixed('USA',ignore_case = TRUE)) ~ 'United States',
    TRUE ~ origin)) %>%
  mutate_if(is.character, factor)


# few european countries
eu <- c(
  "Portugal", "Spain", "France", "Switzerland", "Germany",
  "Austria", "Belgium", "UK", "Netherlands",
  "Denmark", "Norway","Finland","Sweden","Israel","Ireland","Malta","Poland", "Italy",
  "Croatia", "Slovenia", "Hungary", "Slovakia","Greece","Iceland","Luxembourg","Malta","Hungary","Slovakia", "Montenegro","Poland","Lithuania","Czech republic"
)
                                                                                                                                                                                                                                                                                          
# Retrieve the map data
eu.map <- map_data("world", region = eu)
world.map <- map_data("world")
states <- map_data("state")


# Compute the centroid as the mean longitude and latitude
# Used as label coordinate for country's names
eu.lab <- eu.map %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))

states.lab <- states %>%
  summarise(region="United States", long = mean(long), lat = mean(lat))
  
uk_states <- eu.map%>% filter(str_detect('UK', region)) %>%
  group_by(subregion) %>%
  summarize(long=mean(long), lat=mean(lat)) %>%
  rename(region=subregion)
uk_states <- bind_rows(uk_states ,tibble(data.frame(region='England', lat=52, long=0.12)))

eu_countries <- eu.map %>% filter(!str_detect('UK', region)) %>%
  group_by(region) %>%
  summarize(long=mean(long), lat=mean(lat))

hong_kong <- tibble(region="Hong Kong", long=114.177216, lat=22.302711)
geo_loc <- bind_rows(uk_states, eu_countries)

meta_df <- experiment_clean %>% count(origin,insdc, sort=TRUE) %>% rename(region=origin)


hits <- inner_join(geo_loc, meta_df)


world.center  <- world.map %>%  group_by(region) %>%
        summarize(long=mean(long), lat=mean(lat))
world.center <- bind_rows(world.center, uk_states) 
whits <- inner_join(bind_rows(world.center,states.lab, hong_kong), meta_df)

```

```{r}

# Construct URL to ENA for each experiment.
#path <- "http://www.ebi.ac.uk/ena/data/view/"
path <- "https://www.ebi.ac.uk/ena/browser/view/"
experiment_accession <- experiment_clean$experiment_accession
sample_accession <- experiment_clean$sample_accession
experiment_accession =  paste0('<a target=_blank href="',path, experiment_accession,'">', experiment_accession,'</a>' )
sample_accession =  paste0('<a target=_blank href="',path, sample_accession,'">', sample_accession,'</a>' )
experiment_clean$experiment_accession <- experiment_accession
experiment_clean$sample_accession <- sample_accession
```



# Maps {data-icon="fa-map-marker"} 

Row
-----------------------------------------------------------------------

### Total Experiments {.value-box}

```{r}
exp_count <- experiment_clean %>% select(experiment_accession) %>% distinct %>% count() %>% as.numeric()
valueBox(exp_count, icon="ion-ios-flask-outline")
```

### Sequence Release {.value-box}

```{r}
seqrelease_count <- seqrelease %>% select(accession) %>% distinct %>% count() %>% as.numeric()
valueBox(seqrelease_count, icon="ion-ios-flask-outline")
```

### Sequence Updates {.value-box}

```{r}
sequpdate_count <- sequpdate %>% select(accession) %>% distinct %>% count() %>% as.numeric()
valueBox(sequpdate_count, icon="ion-ios-flask-outline")
```


### Submission centers {.value-box}

```{r}
center_count <- experiment_clean %>% select(center_name) %>% distinct %>% count() %>% as.numeric()
valueBox(center_count, icon="ion-ios-people-outline")
```


### Instrument models 

```{r}
models_count <- experiment_clean %>% select(instrument_model) %>% distinct ()%>% count() %>% as.numeric()
valueBox(models_count, icon="ion-paintbucket")
```



### Instrument Platforms 

```{r}
platform_count <- experiment_clean %>% select(instrument_platform) %>% distinct ()%>% count() %>% as.numeric()
valueBox(platform_count, icon="ion-paintbucket")
```


## Row  {.tabset .tabset-fade}

### Europe Experiment


```{r, fig.height=7, fig.width=9}
mybreak=seq(0,nrow(experiment_clean),2000)
mylabel=paste(gsub("000","",mybreak),"K", sep="")
ggplot() +
  geom_map(data=eu.map, map=eu.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  geom_text(aes(x=long, y=lat, label = region), data = geo_loc,  size = 2, hjust = 0.5, family="Palatino") +
  geom_label_repel(data=hits, aes(x=long, y=lat , colour=insdc,label=n), size=2, alpha=.9, box.padding = 0.3, show.legend=T)+
  guides(shape=FALSE) +
  stat_summary_hex(
    data=hits,
    aes(long, lat, z=as.integer(n)),
    fun ="mean",
    alpha = 0.7, 
    bins=90
  ) +
  scale_fill_gradient(high="cyan2", breaks=mybreak, labels=mylabel) + #ggthemes::theme_map() +
  theme_void(base_family = "IBMPlexSans") + 
  labs(x=NULL, y=NULL, fill=NULL, shape=NULL, colour=NULL ) + 
  theme(plot.title= element_text(hjust=.5), 
        legend.position = "bottom",
        legend.text=element_text(size=rel(0.5))) 
```

### World Experiment

```{r, fig.height=7, fig.width=9}
mybreak=seq(0,nrow(experiment_clean),2000)
mylabel=paste(gsub("000","",mybreak),"K", sep="")

ggplot() + 
  geom_map(data=world.map, map=world.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  #geom_label_repel(data=whits, aes(x=long, y=lat ,label=n, colour=insdc, shape = insdc), alpha=.9, box.padding = 0.5)+
  geom_label_repel(data=whits, aes(x=long, y=lat , colour=insdc,label=n), size=2, alpha=.9, box.padding = 0.3, show.legend=T)+
  guides(shape=FALSE) +
  stat_summary_hex(
    data=whits,
    aes(long, lat, z=as.integer(n)),
    fun ="mean",
    alpha = 0.7, 
    bins=90
  ) +
  scale_fill_gradient(high="cyan3", breaks=mybreak, labels=mylabel) + # ggthemes::theme_map() +
  theme_void(base_family = "IBMPlexSans")+ 
  labs(x=NULL, y=NULL, fill=NULL, shape=NULL, colour=NULL) + 
  theme(plot.title= element_text(hjust=.5), 
        legend.position = "bottom",
        legend.text=element_text(size=rel(0.5))) #+ facet_wrap(~insdc)
  


```



### Europe Release/Update

```{r, fig.height=7, fig.width=9}
sequences <- bind_rows(sequpdate %>% mutate(type="Update"), seqrelease %>% mutate(type="Release"))
sequences_df <- sequences %>% count(country,INSDC,type, sort=TRUE) %>% rename(region=country)
seqhits <- inner_join(geo_loc, sequences_df)
wseqhits <- inner_join(bind_rows(world.center,states.lab), sequences_df)


mybreak <- seq(0,nrow(sequences),2000)
mylabel <- paste(gsub("000","K",mybreak),"", sep="")
ggplot() +
  geom_map(data=eu.map, map=eu.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  geom_text(aes(x=long, y=lat, label = region), data = geo_loc,  size = 2, hjust = 0.5, family="Palatino") +
  geom_point(aes(x=long, y=lat), data = seqhits,  size = 1, alpha=.9, colour="steelblue", hjust = 0.5, family="Palatino") +
  geom_label_repel(data=seqhits, aes(x=long, y=lat , colour=INSDC,label=n), size=2, alpha=.9, box.padding = 0.3, show.legend=T)+
  guides(shape=FALSE) +
  #stat_summary_hex(
  #  data=seqhits,
  #  aes(long, lat, z=as.integer(n)),
  # fun ="mean",
  #  alpha = 0.7, 
  #  bins=90
  #) +
  scale_fill_gradient(high="cyan3", breaks=mybreak, labels=mylabel) + #ggthemes::theme_map() +
  theme_void(base_family = "IBMPlexSans") + 
  labs(x=NULL, y=NULL, fill=NULL, shape=NULL, colour=NULL ) + 
  theme(plot.title= element_text(hjust=.5), 
        legend.position = "bottom",
        legend.text=element_text(size=rel(0.5))) + facet_wrap(~type)
```


### World Release/Update

```{r, fig.height=7, fig.width=9}

sequences <- bind_rows(sequpdate %>% mutate(type="Update"), seqrelease %>% mutate(type="Release"))
sequences_df <- sequences %>% count(country,INSDC,type, sort=TRUE) %>% rename(region=country)
seqhits <- inner_join(geo_loc, sequences_df)
wseqhits <- inner_join(bind_rows(world.center,states.lab), sequences_df)

mybreak=seq(0,nrow(sequences),2000)
mylabel=paste(gsub("000","",mybreak),"K", sep="")

ggplot() + 
  geom_map(data=world.map, map=world.map, aes(long, lat, map_id=region), color="white", fill="gray50", size = 0.5, alpha=0.2) + 
  geom_label_repel(data=wseqhits, aes(x=long, y=lat , colour=INSDC,label=n), size=2, alpha=.9, box.padding = 0.3, show.legend=T)+
  geom_point(aes(x=long, y=lat), data = wseqhits,  size = 1, colour="steelblue",alpha=.9, hjust = 0.5, family="Palatino") +
  guides(shape=FALSE) +
  #stat_summary_hex(
  #  data=wseqhits,
  #  aes(long, lat, z=as.integer(n)),
  #  fun ="mean",
  #  alpha = 0.7, 
  #  bins=90
  #) +
  scale_fill_gradient(high="cyan3", breaks=mybreak, labels=mylabel) + # ggthemes::theme_map() +
  theme_void(base_family = "IBMPlexSans")+ 
  labs(x=NULL, y=NULL, fill=NULL, shape=NULL, colour=NULL) + 
  theme(plot.title= element_text(hjust=.5), 
        legend.position = "bottom",
        legend.text=element_text(size=rel(0.5))) + facet_wrap(~type)
  

```


# Experiment {data-icon="fa-calendar"} 

## Row {.tabset .tabset-fade}

### Overall

```{r, fig.height=7, fig.width=9}
source('fileexport.R')

experiment_clean %>% select(first_public, insdc) %>% group_by(first_public, insdc) %>% mutate(count=n()) %>% ungroup() %>% unique() %>% arrange(first_public) %>% group_by(insdc)  %>% mutate(cums=cumsum(count)) %>% ungroup()%>% hchart("line", hcaes(x=first_public, y=cums, group=insdc)) %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
```

### Month

```{r, fig.height=7, fig.width=9}
experiment_clean %>% 
    mutate(monthly=floor_date(first_public, unit="month")) %>% select(monthly, insdc) %>% group_by(monthly, insdc) %>% mutate(count=n()) %>% ungroup() %>% unique() %>% arrange(monthly) %>% group_by(insdc)  %>% mutate(cums=cumsum(count)) %>% ungroup()%>% hchart("line", hcaes(x=monthly, y=cums, group=insdc)) %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))


```


### Week

```{r, fig.height=7, fig.width=9}


experiment_clean %>% 
    mutate(weekly=week(first_public)) %>% select(weekly, insdc) %>% group_by(weekly, insdc) %>% mutate(count=n()) %>% ungroup() %>% unique() %>% arrange(weekly) %>% group_by(insdc)  %>% mutate(cums=cumsum(count)) %>% ungroup()%>% hchart("line", hcaes(x=weekly, y=cums, group=insdc)) %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))



```

# Release/Update {data-icon="fa-calendar"} 

## Row {.tabset .tabset-fade}

### Release
```{r, fig.height=7, fig.width=9}
source('fileexport.R')

sequences %>% filter(str_detect(type, "Release")) %>% select(first_public, INSDC) %>% group_by(first_public, INSDC) %>% mutate(count=n()) %>% ungroup() %>% unique() %>% arrange(first_public) %>% group_by(INSDC)  %>% mutate(cums=cumsum(count)) %>% ungroup()%>%
  mutate(INSDC = ifelse(is.na(INSDC), "None", INSDC)) %>%
  hchart("line", hcaes(x=first_public, y=cums, group=INSDC))%>% hc_subtitle(text="Release") %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
```

### Updates
```{r, fig.height=7, fig.width=9}
source('fileexport.R')

sequences %>% filter(str_detect(type, "Update")) %>% select(first_public, INSDC) %>% group_by(first_public, INSDC) %>% mutate(count=n()) %>% ungroup() %>% unique() %>% arrange(first_public) %>% group_by(INSDC)  %>% mutate(cums=cumsum(count)) %>% ungroup()%>%
  mutate(INSDC = ifelse(is.na(INSDC), "None", INSDC)) %>%
  hchart("line", hcaes(x=first_public, y=cums, group=INSDC))%>% hc_subtitle(text="Updates") %>% 
  hc_exporting(
    enabled = TRUE, 
    filename="simple_column", 
    formAttributes=list(target="_blank"),
    buttons=list(contextButton=list(text="Export", 
                                    theme=list(fill="transparent"),
                                    menuItems=export)))
```



Platforms {data-icon="ion-erlenmeyer-flask"}
=====================================     

## Row {.tabset .tabset-fade}

### Instrument model 

```{r}

experiment_clean %>% count(instrument_model, instrument_platform) %>% mutate(instrument_model = fct_reorder(instrument_model, n)) %>% ggplot(aes(instrument_model, n, fill=instrument_platform)) + geom_col() + coord_flip() 
```

### Instrument platforms 

```{r}

experiment_clean %>% count(instrument_model, instrument_platform, library_layout) %>% mutate(instrument_model = fct_reorder(instrument_model, n)) %>% ggplot(aes(instrument_model, n, fill=instrument_platform)) + geom_col( show.legend = FALSE) + coord_flip() + facet_grid(library_layout~instrument_platform, scales="free_x")
```



Datasets {data-icon="ion-ios-cloud-download-outline"}
=====================================     

## DT::datatable {.tabset .tabset-fade}


### Export Exp. Mapview data


```{r}
library(DT)

whits %>% rename(origin=region, number_experiment=n, submitted_to=insdc) %>%
  datatable( extensions = 'Buttons',  escape=FALSE,
                              options = list(  dom = 'Bfrtip',
                                               buttons = list( list(
                                               extend = 'collection',
                                        buttons = list(list(extend='csv',filename='exp_mapview_data'),                                                                 list(extend='excel', filename = 'exp_mapview_data')),text="Download"
                                            ))))

```

### Basic

```{r}
library(DT)
datatable(experiment_clean, escape=FALSE)
```          

### Advanced

```{r}

datatable(experiment_clean, filter = 'top',  escape=FALSE,
          options = list(   searchHighlight = TRUE, pageLength = 15 ))

```


### Export Experiment

```{r}
library(DT)

experiment_clean %>% datatable( extensions = 'Buttons',  escape=FALSE,
                              options = list(  dom = 'Bfrtip',
                                               buttons = list( list(
                                               extend = 'collection',
                                        buttons = list(list(extend='csv',filename='embl-ebi-covid-submission.csv'),                                                                 list(extend='excel', filename = 'embl-ebi-covid-submission.xlsx')),text="Download"
                                            ))))

```

### Export Sequences releases

```{r}
library(DT)

accession <- sequences$accession
accession =  paste0('<a target=_blank href="',path, accession,'">', accession,'</a>' )
sequences$accession <- accession

sequences %>% datatable( extensions = 'Buttons',  escape=FALSE,
                              options = list(  dom = 'Bfrtip',
                                               buttons = list( list(
                                               extend = 'collection',
                                        buttons = list(list(extend='csv',filename='embl-ebi-covid-sequences.csv'),                                                                 list(extend='excel', filename = 'embl-ebi-covid-sequences.xlsx')),text="Download"
                                            ))))

```
